### 2024-08-08 getting-started.qmd
20240808 - why R, learning R, getting help with R

>>>>

## Learning `R` {#sec-learningR}

Here are a various resources for learning `R`:

- Intro to `R`: <https://www.statmethods.net>
- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
- Learn `R` interactively with `swirl`: <https://swirlstats.com>
- Use the `learnr` package: <https://rstudio.github.io/learnr/>
- Resources for learning `tidyverse`, which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>

## Getting Help with `R` {#sec-gettingHelpR}

If you have `R` questions, you can ask them in a number of places:

- Forums:
    - `Posit`: <https://forum.posit.co>
    - `StackOverflow`: <https://stackoverflow.com/questions/tagged/r>
    - `Reddit`: <https://www.reddit.com/r/rstats/>
- The `R` mailing list: <https://stat.ethz.ch/mailman/listinfo/r-help>

The following article provides additional resources and good guidance: <https://www.r-bloggers.com/where-to-get-help-with-your-r-question/>.

When posting a question on forums or mailing lists, keep a few things in mind:

- Read the posting guidelines before posting!
- Be respectful of other people and their time.
`R` is free software.
People are offering their free time to help.
They are under no obligation to help you.
If you are disrespectful or act like they owe you anything, you will rub people the wrong way and will be less likely to get help.
- Provide a minimal, reproducible example.
Providing a minimal, reproducible example can be crucial for getting a helpful response.
By going to the trouble of creating a minimal reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
Here are guidelines on providing a minimal reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.

### 2024-08-08 intro.qmd
20240527 - statistical programming
20240527 - formatting
20240808 - why R, learning R, getting help with R

How can we use information to make predictions about uncertain events?
This book is about empiricism (basing theories on observed data) and judgment, prediction, and decision making in the context of uncertainty.
The book provides an introduction to modern analytical techniques used to make informed predictions, test theories, and draw conclusions from a given dataset.

@@

@@

Skills in data analysis are highly valuable.
This book includes practical and conceptual tools that build a foundation for critical thinking.
The book aims to help readers evaluate theory in the light of evidence (and vice versa) and to refine decision making in the context of uncertainty.
Readers will learn about the ways that psychological science (and related disciplines) poses questions, formulates hypotheses, designs studies to test those questions, and interprets the findings, collectively with the aim to answer questions, improve decision making, and solve problems.

@@

- Apply empirical inference and appreciate the value it provides over speculative supposition.
- Ask educated questions when confronted with decisions in the face of uncertainty.
- Understand human decision making, including common heuristics and cognitive biases and how to mitigate them analytically.
- Engage in critical thinking about causality, including devising plausible alternative explanations for observed effects.
- Understand causal inference including confounding, causal pathways, and counterfactuals.
- Think empirically about human behavior and performance.
- Describe the strengths and weaknesses of humans versus computers in prediction scenarios.
- Manipulate and summarize datasets.
- Critically evaluate the strengths and limitations of different statistical models and methodologies used in predicting uncertain events, enhancing their understanding of statistical inference and model selection.
- Use various analytical techniques for predicting the outcome of uncertain events, and for uncovering latent causes of patterns in observed data.
- Interpret findings from various statistical approaches and evaluate the accuracy of predictions.
- Engage in iterative problem-solving processes, refining analytical approaches based on feedback and outcomes, and adapting strategies accordingly.
- Communicate statistical findings and analyses in both written and oral formats, demonstrating proficiency in presenting complex information to diverse audiences.
- Make sense of big data.
- Use practical analytical skills that can be applied in future research and job settings.

>>>>

How can we use information to make predictions about uncertain events?
This book is about empiricism (basing theories on observed data) and judgment, prediction, and decision making in the context of uncertainty.
The book provides an introduction to modern analytical techniques used to make informed predictions, test theories, and draw conclusions from a given dataset.
The book leverages the software `R` for providing applied data analysis examples.

@@

## Why R? {#sec-whyR}

The book provides data analysis examples using the statistical analysis software, `R`.
Why `R`?

- `R` is free!
Anyone can use it.
- `R` is open source—it is not a black box.
You can see what is going on "under the hood" and can examine the code for any function or computation you perform.
You can even modify and improve these functions by changing the code, and you can create your own functions.
- `R` is open platform—you can use it on multiple platforms, including Windows, MacOS, and Linux.
- `R` has advanced statistics capabilities.
It was designed for statistical analysis and has strong capabalities for data wrangling.
- `R` has capabilities for state-of-the-art graphics.
It has advanced capabilities for creating statistical graphics.
- `R` is widely used—there is a large community of people who use `R` for data analysis that you can draw upon for help from others.
- `R` analyses are based on code (rather than a graphical user interface), which allows reproducibility—with the same data, code, and setup (platform, `R` version, package versions, etc.), you should get the same answer every time.
There are strong resources available for ensuring your analyses in `R` are reproducible by others [@Gandrud2020].
- Anyone (including you) can contribute `R` packages to the community to improve its functionality.
Statistical experts from all over the world have contributed open source packages to `R` for specialized tasks.
In the chance there is not an `R` package that does what you need to do, you can write a function to perform the task and can contribute it as a package to the community for others to use and improve.
The number of `R` packages contributed to the community is growing at a rapid rate.
As of this writing, over 20,000 packages have been contributed to the [Comprehensive R Archive Network](https://cran.r-project.org/web/packages/) (CRAN).
And many more are stored on publicly available version control repositories like GitHub and GitLab.
Chances are, if there is an analysis you need to do, an `R` package exists to do it.

@@

Skills in statistics, statistical programming, and data analysis are highly valuable.
This book includes practical and conceptual tools that build a foundation for critical thinking.
The book aims to help readers evaluate theory in the light of evidence (and vice versa) and to refine decision making in the context of uncertainty.
Readers will learn about the ways that psychological science (and related disciplines) poses questions, formulates hypotheses, designs studies to test those questions, and interprets the findings, collectively with the aim to answer questions, improve decision making, and solve problems.

@@

- Apply empirical inference and appreciate the value it provides over speculative supposition.
- Ask educated questions when confronted with decisions in the face of uncertainty.
- Understand human decision making, including common heuristics and cognitive biases and how to mitigate them analytically.
- Engage in critical thinking about causality, including devising plausible alternative explanations for observed effects.
- Understand causal inference including confounding, causal pathways, and counterfactuals.
- Think empirically about human behavior and performance.
- Describe the strengths and weaknesses of humans versus computers in prediction scenarios.
- Apply basic skills in statistical programming using `R` to manipulate and summarize datasets and to conduct data analysis.
- Critically evaluate the strengths and limitations of different statistical models and methodologies used in predicting uncertain events, enhancing their understanding of statistical inference and model selection.
- Use various analytical techniques for predicting the outcome of uncertain events, and for uncovering latent causes of patterns in observed data.
- Interpret findings from various statistical approaches and evaluate the accuracy of predictions.
- Engage in iterative problem-solving processes, refining analytical approaches based on feedback and outcomes, and adapting strategies accordingly.
- Communicate statistical findings and analyses in both written and oral formats, demonstrating proficiency in presenting complex information to diverse audiences.
- Make sense of big data.
- Use practical analytical skills that can be applied in future research and job settings.
